Data-Driven Modelling \& \\ Forecasting of Chaotic Systems

### 2022-01-18 projectReport/chapter4.tex
Worked through all tex errors

\begin{Definition}
  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous. Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space. 
\end{Definition}

We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.

%#####################################
\begin{center}
\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
\everypsbox{\scriptstyle}
\begin{psmatrix}
U & U\\%
V & V.
%%%
%  \ncline{1,1}{1,2}^{T} 
%  \ncline{1,1}{2,1} <{\phi}
%   \ncline{2,1}{2,2}^{S}
%  \ncline{1,2}{2,2} > {\phi}
\end{psmatrix}
%#####################################
\end{center}

@@

\begin{Definition}
  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
If the input space $U$ is compact, we refer to the system as compactly driven. 
\end{Definition}

@@

Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)

\begin{Definition}
  [\bf Entire Solution] \label{Dfn_Soln} \rm
  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
\end{Definition}

@@

\begin{Example} \rm \label{ex_halfux}
  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
\end{Example}
% \ednote{Nicely written! Thank you}

@@

\begin{Example}\label{Ex_exp} \rm 
  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$. The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.  
 The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
\end{Example}

@@

\begin{Definition}
  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
\end{Definition}

@@

The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
Now suppose, we for some fixed input 
$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states

\begin{equation} \label{eqn_association}
X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
\end{equation}

@@

It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier. This will lead to a result (for. see, \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}

@@

\begin{Remark}
  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
  For the driven system,  
  $g(u,x)=x^2$ the evolution does not depend on the input at all. Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$. A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$. Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$ 
\end{Remark}

@@

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\begin{proof}
  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
\end{proof}

% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$. To refine the scenario, we consider only systems with unique solutions.

\section{Unique Solution Property}

\begin{Definition}
  [\bf Unique Solution Property] \label{Dfn_usp}\rm
  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
\end{Definition}

One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}. The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.

@@

In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.

% \ednote{How should I explain/qualify RC better here? It's mentioned without really explaining. M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir". B: Thanks!}, 
Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e. there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described. A more detailed presentation may be found in~\cite{jaeger2001echo}.

% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN), a component of the implementation discussed in~\ref{ch5}.

@@

  \begin{Definition}
    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
  \end{Definition}

  If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.

But does such a function $h$ for the above driven system $g$ above exist? Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.

Re-sketching the commutativity diagram \ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below. In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.

@@

When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$. The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).

@@

\begin{Definition}
  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
 The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\} \] is called the inverse-limit space of $(U,T)$.
 The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
\end{Definition}

@@

%To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
In the above example(\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
This is true even for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}

@@

SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity. 
'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).

Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$

The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.

@@

At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications \cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP. The function 
\begin{equation}  \label{eqn_driving}
  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
\end{equation} 
is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP \cite[Th.2]{manjunath2013echo }. 
It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
\begin{equation} \label{eqn_SI_RNN}
  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
  \end{equation}
  when $x_{n-1}$ and $x_n$ are known.

This specific driving function $g$ is used in our implementation and is discussed more completely in chapter~\ref{ch5}

Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property(UAP). 
We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.

\begin{Definition}
  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
  A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
\end{Definition}

@@

\begin{figure}[ht]
  \includegraphics[scale=0.35]{Graphs/_Nonlinearity_Fig_response.eps}
  \centering
\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius. 
Reproduced from \cite{Manju_Nonlinearity}. }
\label{fig:memloss_conttime} 
\end{figure} 
\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately.}

@@

\begin{Theorem}
  [\bf Causal Embedding Theorem (adopted from \cite{Supp})]
  % {\bf (Causal Embedding Theorem.)}
 \label{Thm_CET}
	Let $g$ be a driven system with SI-invertibility and the USP. Let $h$ denote the universal semi-conjugacy and $H_2(\overleftarrow{u}) := (h(r\overleftarrow{u}),h(\overleftarrow{u}))$, where $r$ is the right-shift map. 
 Let $(\widehat{U}_T, \widehat{T})$  be the inverse-limit system of a dynamical system $(U,T)$. 
 Then the restriction of $H_2$ to $\widehat{U}_T$ is a topological semi-conjugacy between the inverse-limit system $(\widehat{U}_T, \widehat{T})$ 
and the induced dynamical system  $(Y_T,G_T)$, i.e., the following diagram commutes
\begin{equation} \label{comm_H_CET}
\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
 \widehat{U}_T & \widehat{U}_T\\%
 Y_T &  Y_T.
 %%%
%  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
%  \ncline{1,2}{2,2} > {H_2}
%  \ncline{2,1}{2,2}^{G_T}
 \end{psmatrix}
 \end{equation}
or in other words, $(Y_T, G_T)$ is a factor of  $(\widehat{U}_T, \widehat{T})$. Further, if $T:U \to U$ is a homeomorphism, 
then $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$, and hence $(Y_T, G_T)$ is conjugate to $(\widehat{U}_T, \widehat{T})$.
\end{Theorem}
{\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}

>>>>

\begin{Definition}
  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP})]\label{Dfn_NDS}\rm
  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps ${\{g_n\}}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous. Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space. 
\end{Definition}

We immediately recall the figure first defined in\eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system. 
See this sketched in the diagram~\ref{UV_conjugacy}

%#####################################
\begin{equation}
  % \begin{center}
    \centering
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
    \everypsbox{\scriptstyle}
    \begin{psmatrix}\label{UV_conjugacy}
    U & U\\%
    V & V.
    %%%
    %  \ncline{1,1}{1,2}^{T} 
    %  \ncline{1,1}{2,1} <{\phi}
    %   \ncline{2,1}{2,2}^{S}
    %  \ncline{1,2}{2,2} > {\phi}
    \end{psmatrix}
    %#####################################
    % \end{center}
\end{equation}

@@

\begin{Definition}
  [\bf Driven, Compactly Driven Dynamical System]\label{Dfn_DDS} \rm
A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
If the input space $U$ is compact, we refer to the system as compactly driven. 
\end{Definition}

@@

Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps ${\{g(u_n, \cdot)\}}_{n\in\mathbb{Z}}$ contained in $X$.
Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)

\begin{Definition}
  [\bf Entire Solution]\label{Dfn_Soln} \rm
  A sequence ${\{x_n\}}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
  \begin{equation}
    g(u_{n-1}, x_{n-1})=x_n \text{ for all }n\in\mathbb{Z}
  \end{equation}

\end{Definition}

@@

\begin{Example}\label{ex_halfux} \rm
  The only solution  ${\{x_n\}}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence, say $u_n=0.5$. 
  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
\end{Example}
% \ednote{Nicely written! Thank you}

@@

\begin{Example}\label{Ex_exp} \rm 
  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$. The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $\lim_{n\to\infty}x_n=0$, $\lim_{n\to{-}\infty}x_n=1$.  
 The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
\end{Example}

@@

\begin{Definition}
  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
\begin{equation}
  X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.  
\end{equation}

The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
\end{Definition}

@@

The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
Now suppose, we for some fixed input 
$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
\begin{equation} \label{eqn_association}
X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
\end{equation}

@@

It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier. This will lead to a result (see~\cite{manjunath2013echo}) that \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}

@@

\begin{Remark}
  [\bf Proof of Example~\ref{Ex_exp}]\label{rem_proofEx} \rm
  For the driven system,  
  $g(u,x)=x^2$ the evolution does not depend on the input at all. Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$. A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$. Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$ 
\end{Remark}

@@

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
{\bf Proof.} May be found in~\cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$. To refine the scenario, we consider only systems with unique solutions. 
\section{Unique Solution Property}

\begin{Definition}
  [\bf Unique Solution Property]\label{Dfn_usp}\rm
  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
\end{Definition}

One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}. The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to~\cite{Manju_ESP, esann2012ids} for further reading.

@@

In the majority of Reservoir Computing (RC) literature- a machine learning approach- where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.

% \ednote{How should I explain/qualify RC better here? It's mentioned without really explaining. M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir". B: Thanks!}, 
Concepts like the ESP,  fading memory~\cite{boyd1985fading} are some of these notions.  
If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e. there is only ever the possibility of a single reachable state at a given instance of time~\cite{jaeger2001echo,Manju_2020}.
% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described. A more detailed presentation may be found in~\cite{jaeger2001echo}.

% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks (RNN), a component of the implementation discussed in~\ref{ch5}.

@@

  \begin{Definition}
    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram~\ref{Scomm_h} commutes for all $v \in U$.
  \end{Definition}

  If the universal semi-conjugacy $h$ exists (i.e. the diagram in~\ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.

But does such a function $h$ for the above driven system $g$ above exist? Whenever $g$ has the USP and $\Psi(u)={\{x_n\}}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above~\ref{Scomm_h}.
Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP~\cite[Lemma 5]{Manju_Nonlinearity}.
Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example~\ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.

Re-sketching the commutativity diagram~\ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below. In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.

@@

When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$. The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter~\ref{ch3}).

@@

\begin{Definition}
  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
 The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by 
 \begin{equation}
  \widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\}  
 \end{equation}
 is called the inverse-limit space of $(U,T)$.
 The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by 
 \begin{equation}
  \widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))  
 \end{equation}
 and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
\end{Definition}

@@

%To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
In the above example (\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
This is true even for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}

@@

SI-invertibility promises that `enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure `enough' information is retained without introducing unwanted complexity. 
`Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).

Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
\begin{equation}
  Y_T:=\{(x_{n-1},x_n): {\{x_k\}}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.  
\end{equation}

The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics (SDD) of the system above.

@@

At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications~\cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.\@The function 
\begin{equation}  \label{eqn_driving}
  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
\end{equation} 
is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP~\cite[Th.2]{manjunath2013echo }. 
It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
\begin{equation} \label{eqn_SI_RNN}
  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
  \end{equation}
  when $x_{n-1}$ and $x_n$ are known.

This specific driving function $g$ is used in our implementation and is discussed more completely in Chapter~\ref{ch5}

Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property (UAP). 
We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project, we refer to~\cite{Manju_Nonlinearity}.

\begin{Definition}
  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
  A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},\ldots$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
\end{Definition}

@@

\begin{figure}[ht]
  \includegraphics[scale=0.35]{Graphs/_Nonlinearity_Fig_response.eps}
  \centering
\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ where $g$ has the USP) and blue ($a=1$, $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.  Reproduced from~\cite{Manju_Nonlinearity}. } 
\label{fig:memloss_conttime} 
\end{figure} 
\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately.}

@@

\begin{Theorem}
  [\bf Causal Embedding Theorem (adopted from~\cite{Supp})]
  % {\bf (Causal Embedding Theorem.)}
\label{Thm_CET}
	Let $g$ be a driven system with SI-invertibility and the USP. Let $h$ denote the universal semi-conjugacy and $H_2(\overleftarrow{u}) := (h(r\overleftarrow{u}),h(\overleftarrow{u}))$, where $r$ is the right-shift map. 
 Let $(\widehat{U}_T, \widehat{T})$  be the inverse-limit system of a dynamical system $(U,T)$. 
 Then the restriction of $H_2$ to $\widehat{U}_T$ is a topological semi-conjugacy between the inverse-limit system $(\widehat{U}_T, \widehat{T})$ 
and the induced dynamical system  $(Y_T,G_T)$, i.e., the following diagram commutes
\begin{equation} \label{comm_H_CET}
\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
 \widehat{U}_T & \widehat{U}_T\\%
 Y_T &  Y_T.
 %%%
%  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
%  \ncline{1,2}{2,2} > {H_2}
%  \ncline{2,1}{2,2}^{G_T}
 \end{psmatrix}
 \end{equation}
or in other words, $(Y_T, G_T)$ is a factor of  $(\widehat{U}_T, \widehat{T})$. Further, if $T:U \to U$ is a homeomorphism, 
then $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$, and hence $(Y_T, G_T)$ is conjugate to $(\widehat{U}_T, \widehat{T})$.
\end{Theorem}
{\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}
